\subsection{\label{sec:history}Historical Context}
As described in the introduction, concepts of cognitive computing lead to the consideration superconducting optoelectronic hardware as a primary candidate for large-scale neural systems capable of general intelligence. Superconducting optoelectronic technology for neural computing resides at the confluence of multiple disciplines. The subject is derived in large part from the foundations of communication theory and computation, as established by Turing \cite{tu1936}, von Neumann \cite{ne1945}, Shannon \cite{sh1948} and others. Yet the nature of information processing in neural systems is not wholly captured by the mathematical analysis formalized to describe serial communication and computation with digital signals. Concepts from neuroscience dating back to Ramon y Cajal \cite{ra1908}, Mountcastle \cite{mo1978}, Edeleman \cite{ed1978}, and many others indicate that clean, serial data streams have much to gain from the network concepts of neural systems with complex spatial and temporal behavior, particularly if one seeks a machine that can think like an intelligent being. If one seeks complexity of performance, one must provide complexity in hardware. If one seeks comprehension in the face of ambiguity, the computation must be able to handle shades of gray. Probability and statistics are inherent to neural systems. 

Conceptually, neural systems contain threads of communication theory, digital logic, probability and statistics. In practice, these systems combine the physics of many devices. The systems envisioned here utilize electrons to compute and photons to communicate. They leverage semiconductors to make light and superconductors to compute. At this confluence of physics and computing, it is possible to conceive of systems with complexity from the chip scale to the globe, employing the logic of neural systems for cognitive information processing across broad reaches of space and time. 

Why do this now? Why this way? Since the inception of the EDVAC, its limitations were anticipated \cite{}. Yet the microelectronic march delayed the consequences by many decades. The scaling first charted by Moore \cite{mo1965} left little impetus for revolutionary concepts. Now that scaling is reaching physical limitations \cite{}, there is an appetite for what new hardware may have to offer. And as computing machines have become deeply integrated in society, we are ready for a sea change in what we ask our machines to do. They can answer any question of fact, but that is no longer all we ask of them. We seek intelligent machines, computers that think, not superficially, but with as deep a wisdom as we can manage to construct. That is why we seek to combine the strengths of superconductors and light. In conjunction, we think, they will lead to the most intelligent machines. To put this discussion in context, we must revisit the origins of computing.

\input{__history_of_digital_computing}

\input{__history_of_neuroscience}

\input{__history_of_ai}

\input{__present_context_and_aim_of_article}