Computers are increasingly being used to aid in information-processing tasks that lack a well-defined mathematical construction and instead require the ability to assimilate information across a broad range of content categories and use that information to make decisions, sometimes with regard to ambiguous criteria. Digital computers were not invented to address these tasks, and attempts to adapt digital hardware for applications requiring general intelligence point to opportunities for significant performance improvements if new hardware can be devised. These opportunities have led to a Cambrian explosion of many types of devices and approaches to beyond-CMOS hardware for artificial intelligence, predominantly in the domain of neural computing. This article reviews the history of both computing and neuroscience in order to establish the context for the present burgeoning of new hardware evolution, and a summary of past and present efforts in neural hardware with particular attention to semiconducting, superconducting, and photonic neural systems is presented. The argument is put forth that a hardware platform combining the strengths of light for communication with superconducting electronics for computation has unique promise to embody the principles of neural information in systems of very large scale. Further, it is argued that a superconducting optoelectronic network may serve as a central cognitive hub in an intelligent system employing specialized modules based on semiconductors, superconductors, or photonics. The specialized modules may perform specialized computations such as numerical analysis, deep learning, or quantum information processing, and the results of those computations can be integrated with the . The analysis indicates the physical possibility to achieve intelligent systems modeled after but potentially far exceeding the capabilities of the human brain.

Based on the spatiotemporal requirements of cognitive activity, the complimentary physics of electronics and photonics, and recent advances in photonic-electronic integration, optoelectronic hardware appears promising and potentially required for large scale artifical coginitive systems. 